Mechanics of gravel rivers: Einstein and Bagnold in modern context

### 2021-09-03 revision/formerPapers/com-piece/scraps/com-piece3.tex
added phd thesis.

\title{On simplification and gravel rivers: Einstein and Bagnold in a modern context}
\author{Kevin Pierce}
\maketitle

\section{Introduction: Gravel bed rivers and mechanistic theories of bedload motion}
%% gravel bed rivers, bedload transport, mechanistic theories, Einstein and Bagnold 
Gravel bed rivers are troubled natural systems, and changes in their morphology are linked to a wide array of ecological and engineering problems.
Since the transport of coarse material as bedload is an important control on channel morphology, bedload transport has been the subject of intense research for over a century. 
Despite this effort, our ability to predict bedload transport rates is inadequate.  
In natural streams, predictions of bedload transport can deviate from measured values by factors of 100 or 1000. 
Prediction is difficult because individual bedload grains move intermittently, rolling, sliding, and bouncing along the river bed \citep{Einstein1950, Bagnold1973}. 
Since this bedload transport process is governed by Newtonian mechanics, it should be fully characterized by exchanges of momentum between the granular and fluid phases, highlighting a mechanical description as an attractive prospect.
However, because bedload transport results from a turbulent flow interacting with granular matter, the full mechanical description is a fusion of two notoriously difficult and tenuously understood subfields of classical physics, meaning many of the required details are missing.
In absence of these details, many researchers have attempted to reverse engineer mechanical theories of bedload transport, and from these efforts, two of them have emerged as important research paradigms in the physics of bedload transport. 
These are the stochastic approach, initiated by \citet{Einstein1937, Einstein1950, Einstein1964} and the deterministic approach, initiated by \citet{Bagnold1956, Bagnold1966, Bagnold1973}.

@@

\section{Bedload fluxes from individual motions: the stochastic paradigm of Einstein}
% Einstein -- bedload flux is individual motions -- individual motions are random -- start to stop transitions -- entrainment, motion, and deposition

@@

\section{Bedload fluxes from collective dissipation: the deterministic paradigm of Bagnold}

%"According to the bagnold hypothesis, the fluid shear stress at the bed equals the threshold value corresponding to incipient motion of particles; the difference %between the total bottom shear stress provides the shear stress in the solid phase. (Ancey2006) Interpreting the excess stress as a result of momentum transfers %between the solid and fluid phases makes it possible to compute the number of particles that can be entrained and maintained in motion, and thereby the solid flow %rate"

Bagnold formed a different simplified concept of bedload transport to understand it with mechanical principles, initatiating the deterministic paradigm of mechanistic bedload transport theory. 
From the physical point of view, Bagnold considered bedload transport in terms of a momentum exchange between solid and liquid phases.
He balanced the average power supplied to the flow by gravity against the average frictional dissipation by moving bedload, and neglected any details of turbulent fluctuations, irregularity in the configuration of stationary grains, or the trajectories of moving particles.
Bagnold's formulation hinges on the application of an average dynamic friction coefficient to moving bedload in order to characterize its energy dissipation. 
His formulation is built upon a key hypothesis: when bedload transport is at its equilibrium value, so it dissipates all available power to move sediment, the tractive force imparted by the moving fluid on the bed surface will match to the threshold for sediment motion. 
This is the so-called Bagnold hypothesis \citep{Engelund1976, Luque1976, Seminara2002, Ancey2006}.
Bagnold's original formulation predicts a bedload transport rate which scales, at high stages of transport, as the average fluid shear stress to the 3/2 power: $q_b \propto \tau^{3/2},$ and his hypothesis has been taken up by generations of researchers, an enduring legacy which is even more interesting in that his bedload transport research seems to be at most a tertiary component of illustrious life's work, with a larger portion of his time devoted to his career as a decorated military commander in both world wars, and his initially recreational hobby of driving cars around the Sahara desert, which eventually gained him status as a preminent desert explorer \citep{Bagnold1988}.

Bagnold's original concept is enlightening for its pragmatic approach \citep{Ashida1972, Bagnold1973, Engelund1976, Luque1976}: his deterministic picture of bedload motion is simplified just enough from the real complexity to make contact with a foundational and reliable concept of classical physics, the conservation of energy. 
At the same time, his highly simplified assumptions and comparisons between experiements and his results have left many open questions for proponents of his ideas.

% Many investigators have built upon and refined Bagnold's concepts while preserving his main idea that bedload transport rates as a consequence of an average energy budget. 
% Early on, investigators came to the conclusion that the details of bedload trajectories are not negligible in a Bagnold-like formulation, and that bedload trajectories cannot be understood without accounting for turbulent lift forces.  
% suggesting a contingence of the bagnold formulation on the details of particle motions ( Luque and Van Beek 1976 )
% For one thing, lifting forces on sediment grains, which are highly variable in accord with fluid turbulence, contribute to the relative amount of time that bedload grains spend in contact with the bed, meaning the dynamic friction coefficient bagnold introduced should scale with fluid lift. 
% accordingly, a large set of bagnold-type treatments focused on including fluid turbulence in-so-far as it affects bedload trajectories, but not in relation to the average power balance which controls, in Bagnold's view, the average number of partilces in motion. 
% These studies effectively contend that bedload is dominated by an average power balance, but the dissipation mechanism within this balance is contingent on fluid turbulence.

For example, Bagnold's total neglect of fluid turbulence and any irregularity between sediment grains left open problems around the degree to which fluid turbulence effects and the differential mobility of grains of varying size can be incorporated into a Bagnold-type framework.  
Likewise, Bagnold's agnostic approach to individual bedload motions within his theory led many investigators to consider how these details can be reconciled with his energy dissipation concept. 
Despite these simplifications, Bagnold's original theory is aptly descriptive of bedload transport under particular conditions: at high transport stages, bedload transport rates often appear to scale as the average shear stress to the $3/2$ power ($q_b \propto \tau^{3/2}$) in accord with his mechanical prediction, at least over flat beds sheared by a steady uniform flow \citep{MeyerPeter1948,Gomez1989,Schmeeckle2014, Elghannay2017}. 
However, Bagnold's theory is troubled at relatively low transport stages and arbitrarily sloping beds.
To describe bedload rates at low transport stages, the empirical components of Bagnold's theory must be set to unphysical values \citep{Engelund1976, Luque1976,  Nelson1995, Nino1998}; and his deterministic theory neglects a key attribute of bedload transport rates at low stages, namely that transport is highly intermittent, so that bedload transport rates exhibit strong fluctuations which Bagnold's theory does not constrain \citep{Bohm2004, Ancey2008}. 
Over arbitrarily sloping beds, there are theoretical arguments that Bagnold's hypothesis should flatly fail, even for relatively small bed slopes transverse to the flow direction \citep{Seminara2002}. 
These problems have motivated intense research within the Bagnold paradigm.

Many researchers have concentrated on computing individual particle trajectories from Newtonian physics, and incorporating these details with Bagnold's average energy dissipation concepts, which should fix the number of particles in motion \citep{Abbott1977, Bridge1984, Wiberg1989, Bridge1992, Nino1998, Bialik2015}. 
Because the details of particle trajectories are highly contingent on fluid turbulence and collisions with the irregular bed surface, this incorporation of bedload trajectories can be viewed as a kind of first order inclusion of turbulent forcing and the random arrangement of bed grains \citep{}, which are key elements of Einstein-type models. 
A difficult attribute of this problem is that energy dissipation due to bedload should happen primarily when bedload is in contact with the bed surface, but the relative fraction of time bedload spends in contact with the bed surface is highly contingent on fluid turbulence, and especially fluid lift, and Bagnold neglected both of these aspects. 
Hence, Bagnold's dynamic friction coefficient should not really be a constant across all transport stages, and is instead expected to depend upon the details of bedload trajectories, which are inherently statistical and incompletely understood \citep{Bialik2015}. 
For one thing, the details of forces imparted by a turbulent flow on bedload grains are incompletely understood \citep{Maxey1986, Schmeeckle2007, Dwivedi2010, Dwivedi2011}, with relatively important terms in the force balance aside from the typical drag and lift \citep{Bialik2015}.
These models of individual bedload motions remain somewhat troubled, usually overpredicting the velocity at which bedload grains travel \citep{Bialik2015}, although the research is steadily generating new understanding.

critical shear stresses have been difficult to consistently define and calculate
and there are indications of difficulty in defining the concept as a consistent consequence of underlying physics

Another deficiency of bagnold's concepts is the failure of the approach over arbitrarily sloping beds
initially, the observation that bagnold's hypothesis did not lead to physical conclusions even for a relatively negligible transverse tilt in the bed surface
motivated the development of an alternate formulation which preserved his average energy balance concept while understanding the number of partilces in motion as a balance of erosion and deposition rates
Parker2003 ammended the bagnold hypothesis by replacing it with an Einstein type hypothesis regarding a balance of erosion and deposition rates 
they assumed relationships between the entrainment and deposition rates and the excess shear stress which, again, do not agree with 
Charru 2004 approached a bagnold-type description with a balance of erosion and deposition rates of particles
which was extended in many ways by the Lajenesse tracer paper and the Charru2006
However it did not make explicit reference to bagnold concepts, and is somehwat of a parallel development with similar conclusions

@@

%bagnold is inadequate at low discharges VanBeek1976, Engelund1976, Nelson1995, where the solid transport rate is not suitably large to dissipate the excess energy available to move bedload grains Engelund1976, Luque1976, implying the bulk particle friction coefficient must be fit to unphysical values, higher than the bulk friction angle of sediment. for example, Nino1998 found a dynamic friction angle of 56.6 degrees was required to describe their experiments at low transport stage, i.e., well above the angle of repose for any natural sediment

@@

%The mean-field treatment applied by Bagnold conflicts with a noticeable feature of sediment transport: its intermittent nature at low flow rates. Ancey2006

@@

% Charru2004 attempts to rectify a bagnold framework with empirical data under viscous flows. They find the areal density of moving particles varies with time, which they link to granular compaction processes. Later on, this framework which borrows bagnoldean concepts with Einsteinian concepts in accord with the Parker2003 extension, is developed to take account of tracer diffusion and to describe the ripple spacing on a sand bed.

@@

%Ancey2008 analyzed energy dissipation by turbulence and by momentum transfers with moving bedload they found at low solid discharges as much as 90 percent of the energy supplied by gravity was disspated by turbulence while at higher discharges, as much as 75 percent can be disspated by momentum transfers with moving bedload so we see a continuous transition between the Einstein regime, in which turbulence is the dominant dissipative mechanism of the fluid flow, and the Bagnold regime, in which friction of moving bedload becomes the dominant mechanism.

@@

\section{
% Synthesis -- Einstein and Bagnold are limit cases, but think like Wolman: both are really missing the real complexity -- the intermediate region
% so it's not surprising that they both fail -- intermediate region means feedbacks with morphology introduce interactions from higher scales
% there is a clear separation between the scales of transport and morphology evolution, meaning the feedback is accentuated
% potential critical phenomena and a mixture of scales driving concordant evolution of bedload transport and morphology
% implying a potentially unsteady transport rate
% the next century should be spent exploring these morphology-transport feedbacks in accord with the ideas of classical morphology

@@

Wolman, in his classic theoretical work on geomorphology of river channels \citep{Wolman1968}, introduced the concept of the competent or effective flow. 
To form this concept he highlighted a balance between the frequency and magnitude of hydrological conditions. 
Extremely high river flows may be incredibly geomorphically active, and they may drastically change river morphology over short timescales, but these flows are rare. 
The real geomorphic work is done by intermediate flow conditions which are both relatively common and relatively large enough to overcome the stability of river channels and effect change in their morphology. 
What I'm saying is, following after Wolman, the real geomorphic work is done by intermediate flow conditions over intermediate time scales, and this intermediate range of conditions falls between the two extremes of pure stochastic and deterministic characteristics, the former mostly insensitive of average characterizations of fluid flow and granular resistance, and the latter mostly contingent on them. 
We contend that Wolman's intermediate range is really untouched by any bedload transport theory, stochastic or deterministic.  
In this range, both varying and averaged attributes of the factors driving and resisting bedload motion matter, and incorporating the varying components may require new theoretical approaches, borrowed from granular physics and fluid dynamics, into a fusion of Bagnold-like and Einstein-like bedload transport theories.
Necessarily, these models should be stochastic, because they have to express stochastic transport at lower rates, but at higher rates their output distributions should sharpen, and their predicted fluxes should scale like tau^3/2, in the absence of morphological or watershed-scale influence. 
To develop these approaches, which incorporate everything we know, or select the part of it we consider relevant, we should stop quibbling. 
Transport is deterministic and stochastic; we need to use mechanics because this problem is too complicated to go about without some guiding principles. 
Accordingly, we should integrate and develop the description, and we should do it fast: salmon stocks are at all-time lows, the climate is changing, and forests are being removed at alarming rates. 
Why should we discuss stochastic versus deterministic when each of them works?

\begin{comment}

Where is this coupling in Einstein or Bagnold, or under what limits might they retain predictive power? 
More generally, where are the deficiencies in our understanding of bedload transport, and why do our predictions in natural channels fail?

To arrive at modern bedload transport models, based on mechanical concepts in the schools of Einstein and Bagnold, we're led to neglect almost everything. 
In effect, Bagnold viewed the river as a machine, where water provides the power to move bedload, and friction on moving sediment dissipates it. 
Bagnold represented the power of the fluid phase by the product of the average (time and space) shear velocity with the average shear stress, and he represented the power dissipated by the sediment phase as the energy per unit time to support a given mass of sediment against frictional drag, which he characterized in an average way.

@@

\end{comment} 
\bibliography{biblio}
\end{document}

>>>>

\title{Mechanics of gravel rivers: Einstein and Bagnold in modern context}
\author{Kevin Pierce}
\maketitle

\section{Introduction}
%% gravel bed rivers, bedload transport, mechanistic theories, Einstein and Bagnold 
Gravel bed rivers are troubled natural systems, and changes in their morphology are linked to a wide array of ecological and engineering problems.
Since the transport of coarse material as bedload is an important control on channel morphology, bedload transport has been the subject of intense research for over a century. 
Despite this effort, our ability to predict bedload transport rates is inadequate.  
In natural streams, predictions of bedload transport can deviate from measured values by factors of 100 or 1000. 
Prediction is difficult because individual bedload grains move intermittently, rolling, sliding, and bouncing along the river bed \citep{Einstein1950, Bagnold1973}. 
Since this bedload transport process is governed by Newtonian mechanics, it should be fully characterized by exchanges of momentum between the granular and fluid phases, highlighting a mechanical description as an attractive prospect.
However, because bedload transport results from a turbulent flow interacting with granular matter, the full mechanical description is a fusion of two notoriously difficult and tenuously understood subfields of classical physics, meaning many of the required details are missing.
In absence of these details, many researchers have attempted to reverse engineer mechanical theories of bedload transport, and from these efforts, two of them have emerged as important research paradigms in the physics of bedload transport. 
These are the stochastic approach, initiated by \citet{Einstein1937, Einstein1950, Einstein1964} and the deterministic approach, initiated by \citet{Bagnold1956, Bagnold1966, Bagnold1973}.

@@

\section{The stochastic paradigm of Einstein}
% Einstein -- bedload flux is individual motions -- individual motions are random -- start to stop transitions -- entrainment, motion, and deposition

@@

\section{The deterministic paradigm of Bagnold}

Bagnold formed a different simplified concept of bedload transport to understand it with mechanical principles, initatiating the deterministic paradigm of mechanistic bedload transport theory. 
From a physical point of view, Bagnold considered bedload transport in terms of a momentum exchange between solid and liquid phases.
He balanced the average power supplied to the flow by gravity against the average frictional dissipation by moving bedload, and neglected any details of turbulent fluctuations, irregularity in the configuration of stationary grains, or the trajectories of moving particles.
Bagnold's formulation hinges on the application of an average dynamic friction coefficient to moving bedload in order to characterize its energy dissipation. 
His formulation is built upon a key hypothesis: when bedload transport is at its equilibrium value, so it dissipates all available power to move sediment, the tractive force imparted by the moving fluid on the bed surface will match the threshold of sediment motion. 
This is the so-called Bagnold hypothesis \citep{Engelund1976, Luque1976, Seminara2002, Ancey2006}.
Bagnold's original formulation predicts a bedload transport rate which scales, at high stages of transport, as the average fluid shear stress to the $3/2$ power: $q_b \propto \tau^{3/2}.$ 
His perspective on bedload transport has been taken up by generations of river scientists, and this enduring legacy is even more interesting because his research seems to be at most a tertiary component of his life's work, with a much larger portion of his time devoted to his decorated career as a military commander in both world wars, and his recreational hobby of driving cars around the Sahara desert, which eventually earned him status as a preminent desert explorer \citep{Bagnold1988}.

Bagnold's original concept is enlightening for its pragmatic approach \citep{Ashida1972, Bagnold1973, Engelund1976, Luque1976}. 
His deterministic concept of bedload motion is simplified just enough from the real complexity to make contact with the foundational and reliable mechanical principle of conservation of energy. 
His theory has been influential for a clear ability to describe bedload transport from mechanical concepts at relatively high transport stages, since it predicts a transport rate which scales as the average shear stress to the $3/2$ power, in accord with a large set of classic experiments and contemporary numerical simulations \citep{MeyerPeter1948,Gomez1989,Schmeeckle2014, Elghannay2017}.
At the same time, his highly simplified assumptions, and various comparisons between experiments and his theoretical results, have left many open questions for proponents of his ideas. 
There is at least one criticism of Bagnold's hypothesis that the fluid should be just competent to move sediment on the bed surface under equilibrium bedload transport \citep{Seminara2002}.
In particular, Bagnold's concept of bedload transport ignores the details of bedload trajectories, motivating many efforts to reconcile his average energy budget concepts with the dynamics of individual bedload motions. 
Finally, his theory is apparently inadequate at low discharges \citep{Engelund1976, Luque1976, Francis1977, Ancey2008}, and this realization has supported a great deal of research into why this is so, and how this problem can be rectified.

However, the conceptual simplicity of Bagnold's ideas and the essential correctness of his results at high transport rates is striking, so many investigators have worked to refine the details of his formulation while preserving his main concept that bedload transport rates derive from an average budget between gravitational energy and frictional dissipation. 
The time spent in suspension by bedload grains between successive collisons, which is highly variable in accord with fluid turbulence and collisions with the bed surface \citep{Bialik2015}, contribute to the relative amount of time that bedload grains spend in contact with the bed, meaning the dynamic friction coefficient Bagnold introduced should scale with the characteristics of fluid turbulence, or equivalently with the details of bedload trajectories. 
Accordingly, many investigators have tried to understand the turbulent basis of bedload trajectories, and to incorporate these details with Bagnold's average energy budget concepts, which should fix the number of particles in motion \citep{Abbott1977, Bridge1984, Wiberg1989, Bridge1992, Nino1998}.
This research is a perplexing juxtaposition of average and turbulent energy considerations, and effectively it contends that the number of particles in motion results from an average power balance, while the dissipation mechanism which limits the number of particles in motion is contingent on fluid turbulence and granular collisions. 
A shortcoming of this research is that the details of forces imparted by a turbulent flow on bedload grains are incompletely understood \citep{Maxey1986, Schmeeckle2007, Dwivedi2010, Dwivedi2011}, with relatively important terms entering the force balance apart from the typical drag and lift \citep{Bialik2015}.
Accordingly, these Bagnold-type models incorporating the Newtonian physics of individual bedload motions remain somewhat troubled, usually overpredicting the velocity at which individual grains travel, and therefore also overpredicting bedload transport rates \citep{Bridge1984, Bialik2015}. 
These overpredictions have been attributed to a lack of inclusion of collisions between moving grains \citep{Lee2002}, and a lack of understanding of stresses imparted by moving grains on the stationary bed \citep{Nino1998}, so both of these topics deserve further research attention \citep{Bialik2015}.

One criticism of Bagnold's hypothesis is that it fails under even a relatively small transverse (cross-stream) tilt in the bed surface \citep{Seminara2002}. 
In particular, under such conditions it is shown that no bedload transport rate, no matter how large, is sufficient to reduce the fluid shear stress at the bed to the critical value.
This undermines the Bagnold hypothesis, even under nearly horizontal beds, as it is not robust to realistic deviations from ideal conditions. 
\citet{Parker2003} ammended the Bagnold hypothesis by replacing it with an Einstein-type concept regarding the balance of erosion and deposition rates, although they assumed somewhat ad hoc relationships to develop this connection. 
Nevertheless, these efforts key into a parallel stream of research which borrows many concepts from Bagnold, where balances between erosion and deposition rates, parameterized by average characteristics of the fluid flow, are conjoined to form bedload transport rates with same scaling patterns as Bagnold's original theories, although they do not defer to the energy balance concept to the same degree as Bagnold \citep{Charru2004, Charru2006, Lajenesse2010, Lajenesse2018}.

At low transport stages, Bagnold's theory fails to characterize bedload transport.
In relation to this, one observation is that bedload transport rates become highly fluctuating at relatively low mean discharges, meaning mean values increasingly represent a simplified characteristic of the bedload flux, and statistical representations are really closer to the underlying physics \citep{Ancey2008}.
However, the deepest issue with Bagnold under these conditions is that solid transport rates is not suitably large to dissipate the excess energy available to move bedload grains, meaning Bagold-type models do not fit experimental data without calibrating bulk particle friction coefficients to unphysical values \citep{Engelund1976, Luque1976, Nelson1995, Nino1998}.
For example, \citet{Nino1998} found a dynamic friction angle of $56.6$ degrees was required to describe their experiments at low transport stage, a value well above the angle of repose for any natural sediment.
This failure of Bagnold at low rates is apparently consequent from his underyling assumption that bedload transport is always a dominant energy dissipation mechanism of the flow. 
\citet{Ancey2008} analyzed energy dissipation by turbulence and by momentum transfers with moving bedload.
They found at low solid discharges, as much as $90$ percent of the energy supplied by gravity was disspated by turbulence (so $10$ percent by bedload), while at high solid discharges, a much as $75$ percent was dissipated by momentum transfers with moving bedload (so $25$ percent by turbulence). 
This data suggests that Bagnold's assumptions are only close to the physical reality of bedload transport at higher discharges, and at lower discharges, a different description is required.

@@

@@

@@

@@

@@

\section{Synthesis}
% Synthesis -- Einstein and Bagnold are limit cases, but think like Wolman: both are really missing the real complexity -- the intermediate region
% so it's not surprising that they both fail -- intermediate region means feedbacks with morphology introduce interactions from higher scales
% there is a clear separation between the scales of transport and morphology evolution, meaning the feedback is accentuated
% potential critical phenomena and a mixture of scales driving concordant evolution of bedload transport and morphology
% implying a potentially unsteady transport rate
% the next century should be spent exploring these morphology-transport feedbacks in accord with the ideas of classical morphology

@@

Wolman, in his classic theoretical work on geomorphology of river channels \citep{Wolman1968}, introduced the concept of the competent or effective flow. 
To form this concept he highlighted a balance between the frequency and magnitude of hydrological conditions. 
Extremely high river flows may be incredibly geomorphically active, and they may drastically change river morphology over short timescales, but these flows are rare. 
The real geomorphic work is done by intermediate flow conditions which are both relatively common and relatively large enough to overcome the stability of river channels and effect change in their morphology. 
What I'm saying is, following after Wolman, the real geomorphic work is done by intermediate flow conditions over intermediate time scales, and this intermediate range of conditions falls between the two extremes of pure stochastic and deterministic characteristics, the former mostly insensitive of average characterizations of fluid flow and granular resistance, and the latter mostly contingent on them. 
We contend that Wolman's intermediate range is really untouched by any bedload transport theory, stochastic or deterministic.  
In this range, both varying and averaged attributes of the factors driving and resisting bedload motion matter, and incorporating the varying components may require new theoretical approaches, borrowed from granular physics and fluid dynamics, into a fusion of Bagnold-like and Einstein-like bedload transport theories.
Necessarily, these models should be stochastic, because they have to express stochastic transport at lower rates, but at higher rates their output distributions should sharpen, and their predicted fluxes should scale like $\tau^3/2$, in the absence of morphological or watershed-scale influence. 
To develop these approaches, which incorporate everything we know, or select the part of it we consider relevant, we should stop quibbling. 
Transport is deterministic and stochastic; we need to use mechanics because this problem is too complicated to go about without some guiding principles. 
Accordingly, we should integrate and develop the description, and we should do it fast: salmon stocks are at all-time lows, the climate is changing, and forests are being removed at alarming rates. 
Why should we discuss stochastic versus deterministic when each of them works?

%\begin{comment} 
%
%
%Where is this coupling in Einstein or Bagnold, or under what limits might they retain predictive power? 
%More generally, where are the deficiencies in our understanding of bedload transport, and why do our predictions in natural channels fail?%
%
%To arrive at modern bedload transport models, based on mechanical concepts in the schools of Einstein and Bagnold, we're led to neglect almost everything. 
%In effect, Bagnold viewed the river as a machine, where water provides the power to move bedload, and friction on moving sediment dissipates it. 
%Bagnold represented the power of the fluid phase by the product of the average (time and space) shear velocity with the average shear stress, and he represented the %power dissipated by the sediment phase as the energy per unit time to support a given mass of sediment against frictional drag, which he characterized in an average %way.

%\end{comment}

@@

%\end{comment} 
\bibliography{biblio}
\end{document}
